---
title: "Wave 0 Data Cleaning"
author: "Taylor Guthrie"
date: "2026-02-20"
output: html_document
---


> **Repository/privacy note**
>
> This repository should contain code and documentation only. Raw survey data,
> linkage files, editable audit files, receipts, processed participant-level
> outputs, and rendered HTML are intentionally kept outside version control.
> Some generated audit files contain direct identifiers (including names and
> Qualtrics ResponseIds) and must remain in approved secure project storage.
>
> Researcher/test names and participant-specific name corrections are loaded
> from a private configuration file and are intentionally not embedded in this
> public script.
>
> The script is designed as a staged workflow. When a new editable audit file is
> created, execution stops so that the file can be reviewed manually before the
> corresponding decisions are applied on the next run.








```{r}
###############################################################################
#Libraries and data import
###############################################################################

suppressPackageStartupMessages({
  library(haven)
  library(dplyr)
  library(stringr)
  library(stringi)
  library(tibble)
  library(readxl)
  library(readr)
  library(tidyr)
  library(purrr)
})

# ---- repository-safe configuration ----
# Keep the private data root outside the Git repository.
# Example (Windows, set in ~/.Renviron):
#   GUTS_DATA_ROOT=X:/secure/project/data
dir0 <- Sys.getenv("GUTS_DATA_ROOT", unset = "")
if (!nzchar(dir0)) {
  stop(
    "GUTS_DATA_ROOT is not set. Point it to the private project data root ",
    "before running this pipeline."
  )
}
dir0 <- normalizePath(dir0, winslash = "/", mustWork = TRUE)

# Optional private configuration for researcher/test names and participant-
# specific linkage corrections. Keep this file outside the repository.
# It may define:
#   researcher_names
#   linkage_name_patches
private_config <- Sys.getenv("GUTS_PRIVATE_CONFIG", unset = "")
if (nzchar(private_config)) {
  if (!file.exists(private_config)) {
    stop("GUTS_PRIVATE_CONFIG does not exist: ", private_config)
  }
  source(private_config, local = TRUE)
}

# Avoid embedding participant-level console output in rendered HTML.
knitr::opts_chunk$set(
  echo = TRUE,
  results = "hide",
  message = FALSE,
  warning = FALSE
)

# ---- raw file paths (Wave-specific) ----
paths <- list(
  ic_wave0   = file.path(dir0, "behavioral/T0/data_raw/T0+GUTS+WP2+Informed+Consent_January+15,+2026_04.03.sav"),
  demo_wave0 = file.path(dir0, "behavioral/T0/data_raw/T0+GUTS+WP2+Sensitive+Data_January+15,+2026_04.06.sav"),
  survey_wave0 = file.path(dir0, "behavioral/T0/data_raw/T0+GUTS+WP2+Survey_January+15,+2026_04.07.sav"),
  linkage    = file.path(dir0, "T1_Behavioral/Hackathon/GUTS_WP2_Linkage.xlsx")
)

read_sav_tagged <- function(path, label) {
  stopifnot(file.exists(path))
  df <- read_sav(path) %>% as_tibble()
  message(label, ": ", nrow(df), " rows, ", ncol(df), " cols")
  df
}

# ---- read everything ----
ic_wave0       <- read_sav_tagged(paths$ic_wave0, "ic_wave0")
demo_raw_wave0 <- read_sav_tagged(paths$demo_wave0, "demo_raw_wave0")
survey_wave0   <- read_sav_tagged(paths$survey_wave0, "survey_wave0")

linkage <- read_excel(paths$linkage) %>% as_tibble()
message("linkage: ", nrow(linkage), " rows, ", ncol(linkage), " cols")


```


```{r}
###############################################################################
#Helper functions: names and dates
###############################################################################
# These helpers standardize names and timestamps so comparisons behave
# consistently across SPSS imports and Qualtrics exports.
#
# clean_name():
#   - produces a lowercase, whitespace-squished, NFC-normalized string
#   - used as the stable matching key for identity operations
#
# title_case_name():
#   - title-cases names for later matching to linkage (human-readable form)
#
# fix_tussenvoegsels():
#   - normalizes common Dutch tussenvoegsels after title-casing
#   - improves linkage later when matching to the master linkage file
#
# parse_dt_any():
#   - parses timestamp formats observed across files
#   - returns POSIXct in UTC for consistent comparisons
###############################################################################

clean_name <- function(x) {
  x %>%
    as.character() %>%
    str_squish() %>%
    str_to_lower() %>%
    stringi::stri_trans_nfc()
}

title_case_name <- function(x) {
  x %>%
    as.character() %>%
    str_squish() %>%
    str_to_title()
}

fix_tussenvoegsels <- function(x) {
  x %>%
    str_replace_all("\\sVan\\s", " van ") %>%
    str_replace_all("\\sDer\\s", " der ") %>%
    str_replace_all("\\sDe\\s",  " de ") %>%
    str_replace_all("\\sDen\\s", " den ") %>%
    str_replace_all("\\sTer\\s", " ter ") %>%
    str_replace_all("\\sLe\\s",  " le ") %>%
    str_replace_all("\\sTot\\s", " tot ")
}

parse_dt_any <- function(x) {
  x <- as.character(x)
  out1 <- suppressWarnings(as.POSIXct(x, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"))
  out2 <- suppressWarnings(as.POSIXct(x, format = "%d-%b-%Y %H:%M:%S", tz = "UTC"))
  out  <- ifelse(!is.na(out1), out1, out2)
  as.POSIXct(out, tz = "UTC", origin = "1970-01-01")
}

convert_dates <- function(df) {
  df %>%
    mutate(
      StartDate = parse_dt_any(StartDate),
      EndDate   = parse_dt_any(EndDate)
    )
}

# ---- apply date parsing (Wave 0 already does this; now standardized) ----
ic_wave0        <- convert_dates(ic_wave0)
demo_raw_wave0  <- convert_dates(demo_raw_wave0)
survey_wave0    <- convert_dates(survey_wave0)

```



```{r}
###############################################################################
#Audit IO helpers (WRITE -> EDIT -> READ -> APPLY)
###############################################################################
# Each audit step writes a small CSV to audit_dir. You edit it once,
# re-import it, and we apply the decisions before the next step.
###############################################################################

audit_dir <- file.path(dir0, "behavioral", "T0", "audit_outputs")
if (!dir.exists(audit_dir)) dir.create(audit_dir, recursive = TRUE)

write_audit_csv <- function(df, filename) {
  out_path <- file.path(audit_dir, filename)
  write.csv2(df, file = out_path, row.names = FALSE, na = "")
  message("Wrote audit file: ", out_path)
  invisible(out_path)
}

read_audit_csv2 <- function(filename) {
  in_path <- file.path(audit_dir, filename)
  stopifnot(file.exists(in_path))
  read.csv2(in_path, stringsAsFactors = FALSE)
}

# Editable audit files are deliberately different from read-only reports.
# On first creation, stop execution if there are rows to review. On later runs,
# preserve the edited file rather than overwriting the reviewer's decisions.
write_review_csv <- function(df, filename) {
  out_path <- file.path(audit_dir, filename)

  if (file.exists(out_path)) {
    message("Using existing review file: ", out_path)
    return(invisible(out_path))
  }

  write.csv2(df, file = out_path, row.names = FALSE, na = "")
  message("Wrote review file: ", out_path)

  if (nrow(df) > 0) {
    stop(
      "Manual review required. Edit the decision columns in ", filename,
      " in secure project storage, then re-run this audit step."
    )
  }

  invisible(out_path)
}

```



```{r}
###############################################################################
# Audit Step A (Wave 0): Identify and remove test responses
#
# Goal
# Remove researcher/test activity as early as possible. In Wave 0, the Informed
# Consent (IC) file is the only place where participant names exist. The IC
# Qualtrics ResponseId is treated as the canonical participant identifier (PPN)
# that should propagate downstream into Sensitive Data (demo) and Survey.
#
# Strategy
# Audit Step A has four stages:
#   A1. Construct the full IC universe for the wave and compute conservative
#       test flags (FLAG ONLY).
#   A2. Write one self-contained review file for flagged IC ResponseIds.
#   A3. Re-import the edited review file and drop ResponseIds marked as tests.
#   A4. Propagate the drops globally by removing those PPNs from downstream
#       datasets (demo + survey).
#
# Outputs
#   - A_test_drop_decisions.csv   (EDIT THIS FILE)
#
# Notes
# - Because names do not exist outside IC in Wave 0, we cannot do the Wave 1
#   “FC name → IC name → PPN” tracing. Step A therefore relies on IC date/name
#   flags plus human review in the audit CSV.
###############################################################################


###############################################################################
# A1. Build the IC universe and compute conservative test flags (flag only)
###############################################################################

# ---- WAVE 0 SPECIFIC: official launch time ----
# Entries before this time are strong candidates for piloting/testing.
real_start <- as.POSIXct("2024-10-11 00:00:00", tz = "UTC")

# ---- WAVE 0 SPECIFIC: researcher/test names (best effort) ----
# These are only used to FLAG; final keep/drop is decided in the audit CSV.
# Keep this list broad; the audit file is the true decision layer.
if (!exists("researcher_names", inherits = FALSE)) {
  researcher_names <- character(0)
}

researcher_names_clean <- clean_name(researcher_names)

# Standardize IC fields for consistent auditing
ic_combined <- ic_wave0 %>%
  rename(Name = ic_name) %>%
  mutate(
    StartDate_posix = parse_dt_any(StartDate),
    EndDate_posix   = parse_dt_any(EndDate),
    ResponseId      = as.character(ResponseId),
    ppn             = as.character(ppn),

    # Title-case for human readability + tussenvoegsels cleanup
    Name_title      = fix_tussenvoegsels(str_to_title(as.character(Name))),
    Name_clean      = clean_name(Name_title),

    # Flag only
    flag_test_by_date = !is.na(StartDate_posix) & StartDate_posix < real_start,

    # Exact match list or literal 'test' token in the name (flag only)
    flag_test_by_name = ifelse(
      is.na(Name_clean) | Name_clean == "",
      FALSE,
      (Name_clean %in% researcher_names_clean) | str_detect(Name_clean, "\\btest\\b")
    )
  )

# IC ledger for auditing: full IC universe with flags but no decisions applied
ic_ledger_raw <- ic_combined %>%
  select(
    Name, Name_title, Name_clean,
    ResponseId, ppn,
    StartDate, EndDate, StartDate_posix, EndDate_posix,
    flag_test_by_date, flag_test_by_name
  )


###############################################################################
# A2. WRITE: export one review file for flagged IC ResponseIds (edit decisions)
###############################################################################
# Editable fields:
#   - keep
#   - drop_reason
#   - notes
#
# Do not edit:
#   - ResponseId
#   - Name_clean
###############################################################################

test_rows_review <- ic_ledger_raw %>%
  mutate(
    flag_test = flag_test_by_date | flag_test_by_name,
    test_reason = case_when(
      flag_test_by_date & flag_test_by_name ~ "prelaunch + researcher_name_or_test",
      flag_test_by_date ~ "prelaunch",
      flag_test_by_name ~ "researcher_name_or_test",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(flag_test) %>%
  arrange(StartDate_posix) %>%
  select(
    Name, Name_title, Name_clean,
    ResponseId, ppn,
    StartDate_posix, EndDate_posix,
    flag_test_by_date, flag_test_by_name, test_reason
  )

test_drop_decisions_template <- test_rows_review %>%
  transmute(
    # identity keys (do not edit)
    ResponseId,
    Name_title,
    Name_clean,

    # audit context
    StartDate_posix,
    EndDate_posix,
    flag_test_by_date,
    flag_test_by_name,
    test_reason,

    # editable decisions
    keep = FALSE,
    drop_reason = test_reason,
    notes = ""
  )

write_review_csv(test_drop_decisions_template, "A_test_drop_decisions.csv")


###############################################################################
# A3. READ + APPLY: re-import the edited decisions and drop marked tests
###############################################################################

test_drop_decisions <- read_audit_csv2("A_test_drop_decisions.csv")

# Guardrails: prevent deletions/reordering of the audit universe
stopifnot(all(test_drop_decisions$ResponseId %in% test_drop_decisions_template$ResponseId))
stopifnot(all(test_drop_decisions$ResponseId == test_drop_decisions_template$ResponseId))

drop_test_responseids <- test_drop_decisions %>%
  filter(keep == FALSE) %>%
  pull(ResponseId) %>%
  unique()

# Apply drops to IC (create post Step A working IC objects)
ic_combined_working <- ic_combined %>%
  filter(!(ResponseId %in% drop_test_responseids))

ic_ledger_working <- ic_ledger_raw %>%
  filter(!(ResponseId %in% drop_test_responseids))


#rename ppn column in survey file to match structure of other datasets 
survey_wave0 <- survey_wave0 %>%
  rename(ppn = ppn_) %>%
  mutate(ppn = as.character(ppn)) %>%
  filter(!(ppn %in% drop_test_responseids))


###############################################################################
# A4. APPLY GLOBALLY: remove dropped test PPNs from downstream datasets
###############################################################################
# In Wave 0, downstream datasets store IC ResponseId as `ppn`. Therefore, any
# IC ResponseId dropped in Step A should be removed from demo and survey via
# their `ppn` field.
#
# Implementation note:
# We overwrite ic_combined and ic_ledger_raw to keep later steps clean.
###############################################################################

# 1) IC: overwrite the main objects used downstream in the script
ic_combined <- ic_combined_working
ic_ledger_raw <- ic_ledger_working


# 2) Downstream datasets: drop by PPN
demo_raw_wave0 <- demo_raw_wave0 %>%
  mutate(ppn = as.character(ppn)) %>%
  filter(!(ppn %in% drop_test_responseids))

survey_wave0 <- survey_wave0 %>%
  mutate(ppn = as.character(ppn)) %>%
  filter(!(ppn %in% drop_test_responseids))


###############################################################################
# Post Step A standardization: add parsed timestamps to downstream datasets
###############################################################################
# Not a new audit step. It prepares demo + survey for later audit steps that
# use time matching, completion checks, or duplicate resolution.
###############################################################################

demo_raw_wave0 <- demo_raw_wave0 %>%
  mutate(
    ppn = as.character(ppn),
    StartDate_posix = parse_dt_any(StartDate),
    EndDate_posix   = parse_dt_any(EndDate)
  )

survey_wave0 <- survey_wave0 %>%
  mutate(
    ppn = as.character(ppn),
    StartDate_posix = parse_dt_any(StartDate),
    EndDate_posix   = parse_dt_any(EndDate)
  )

```



```{r}
###############################################################################
# AUDIT STEP B (Wave 0): missing ppn in main survey chain (Survey -> Demo -> IC)
#   Output: ONE review file + receipts written to receipts/ subfolder
###############################################################################
# Prereqs expected in environment (from earlier sections):
#   - audit_dir, write_audit_csv(), read_audit_csv2()
#   - parse_dt_any(), clean_name(), fix_tussenvoegsels()
#   - ic_combined (or ic_combined_working after Step A test drops)
#   - demo_raw_wave0, survey_wave0
###############################################################################

# -------------------------
# 0) Parameters + folders
# -------------------------
tolerance_secs <- 2

receipts_dir <- file.path(audit_dir, "receipts")
if (!dir.exists(receipts_dir)) dir.create(receipts_dir, recursive = TRUE)

write_receipt_csv <- function(df, filename) {
  out_path <- file.path(receipts_dir, filename)
  write.csv2(df, file = out_path, row.names = FALSE, na = "")
  message("Wrote receipt file: ", out_path)
  invisible(out_path)
}

# -------------------------
# 1) Local helpers (self-contained for this section)
# -------------------------
is_missing_ppn <- function(x) {
  x2 <- trimws(as.character(x))
  is.na(x2) | x2 == ""
}

has_missing_ppn <- function(df, col = "ppn") {
  if (!col %in% names(df)) return(FALSE)
  any(is_missing_ppn(df[[col]]))
}

# Strict time match helper:
# Returns best match index + diagnostics, or NA if nothing within tolerance.
time_match_one <- function(target_time, candidate_times, tolerance_secs) {
  if (is.na(target_time) || length(candidate_times) == 0) {
    return(list(best_index = NA_integer_, best_abs_diff = NA_real_, n_within_tol = 0L))
  }
  diffs_abs <- abs(as.numeric(difftime(candidate_times, target_time, units = "secs")))
  n_within <- sum(diffs_abs <= tolerance_secs, na.rm = TRUE)
  if (all(is.na(diffs_abs))) {
    return(list(best_index = NA_integer_, best_abs_diff = NA_real_, n_within_tol = as.integer(n_within)))
  }
  best_idx <- which.min(diffs_abs)
  best_abs <- diffs_abs[best_idx]
  if (is.na(best_abs) || best_abs > tolerance_secs) {
    return(list(best_index = NA_integer_, best_abs_diff = best_abs, n_within_tol = as.integer(n_within)))
  }
  list(best_index = as.integer(best_idx), best_abs_diff = as.numeric(best_abs), n_within_tol = as.integer(n_within))
}

# -------------------------
# 2) Standardize timestamps + key fields used for matching
#    (do not assume earlier sections created *_posix)
# -------------------------

# Choose IC source post Step A test drops if present
ic_source <- if (exists("ic_combined_working")) ic_combined_working else ic_combined

# IC provides canonical PPN as ResponseId (ppn may or may not exist)
ic_B <- ic_source %>%
  dplyr::mutate(
    ic_row_id = dplyr::row_number(),
    ResponseId = trimws(as.character(ResponseId)),
    ppn_ic = ResponseId,
    StartDate_posix = parse_dt_any(StartDate),
    EndDate_posix   = parse_dt_any(EndDate)
  )

demo_B <- demo_raw_wave0 %>%
  dplyr::mutate(
    row_id = dplyr::row_number(),
    ppn = trimws(as.character(ppn)),
    StartDate_posix = parse_dt_any(StartDate),
    EndDate_posix   = parse_dt_any(EndDate)
  )

survey_B <- survey_wave0 %>%
  dplyr::mutate(
    row_id = dplyr::row_number(),
    ppn = trimws(as.character(ppn)),
    StartDate_posix = parse_dt_any(StartDate),
    EndDate_posix   = parse_dt_any(EndDate)
  )

# -------------------------
# 3) Preflight: do we need any repairs?
# -------------------------
need_fix_survey <- has_missing_ppn(survey_B, "ppn")
need_fix_demo   <- has_missing_ppn(demo_B, "ppn")

# Always create objects so downstream code never fails
survey_review_tbl <- tibble::tibble()
demo_review_tbl   <- tibble::tibble()

# -------------------------
# 4) Build Survey suggestions (closest to “end of chain”)
#    Match Survey StartDate to Demo EndDate, then inherit Demo ppn.
# -------------------------
if (need_fix_survey) {

  survey_missing <- survey_B %>%
    dplyr::filter(is_missing_ppn(ppn)) %>%
    dplyr::transmute(
      component = "survey",
      row_id,
      StartDate_posix,
      EndDate_posix,
      match_target_time = StartDate_posix
    )

  survey_review_tbl <- survey_missing %>%
    dplyr::rowwise() %>%
    dplyr::mutate(
      .m = list(time_match_one(match_target_time, demo_B$EndDate_posix, tolerance_secs)),
      demo_row_id = .m$best_index,
      demo_best_abs_diff = .m$best_abs_diff,
      demo_n_within_tol = .m$n_within_tol,
      demo_within_tol = !is.na(demo_row_id),

      demo_StartDate_posix = if (demo_within_tol) demo_B$StartDate_posix[demo_row_id] else as.POSIXct(NA),
      demo_EndDate_posix   = if (demo_within_tol) demo_B$EndDate_posix[demo_row_id] else as.POSIXct(NA),
      demo_ppn_raw         = if (demo_within_tol) demo_B$ppn[demo_row_id] else NA_character_,

      # signed delta for audit (demo end - survey start)
      demo_time_diff_secs = if (demo_within_tol)
        as.numeric(difftime(demo_EndDate_posix, match_target_time, units = "secs"))
      else NA_real_
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      suggested_ppn = dplyr::case_when(
        demo_within_tol & !is.na(demo_ppn_raw) & trimws(demo_ppn_raw) != "" ~ demo_ppn_raw,
        TRUE ~ NA_character_
      ),
      suggestion_status = dplyr::case_when(
        is.na(StartDate_posix) ~ "missing_survey_startdate",
        is.na(demo_row_id) ~ "no_demo_time_match",
        is.na(demo_ppn_raw) | trimws(demo_ppn_raw) == "" ~ "matched_demo_but_demo_ppn_missing",
        TRUE ~ "inherited_from_demo"
      ),
      recommend_apply = !is.na(suggested_ppn) & trimws(suggested_ppn) != ""
    ) %>%
    dplyr::select(
      component, row_id,
      StartDate_posix, EndDate_posix,

      # Demo evidence
      demo_row_id,
      demo_StartDate_posix, demo_EndDate_posix,
      demo_ppn_raw,
      demo_time_diff_secs,
      demo_n_within_tol,

      # suggestion
      suggested_ppn,
      suggestion_status,
      recommend_apply
    )
}

# Build a temporary Survey ppn map from suggestions (used only for diagnostics if needed)
survey_suggested_map <- survey_review_tbl %>%
  dplyr::filter(!is.na(suggested_ppn) & trimws(suggested_ppn) != "") %>%
  dplyr::transmute(
    survey_row_id = row_id,
    survey_suggested_ppn = suggested_ppn,
    survey_suggestion_status = suggestion_status
  )

# Create a Survey working proposal (NOT yet applied; used only if you want later steps)
survey_working_proposal <- survey_B %>%
  dplyr::left_join(survey_suggested_map, by = c("row_id" = "survey_row_id")) %>%
  dplyr::mutate(
    ppn_proposed = dplyr::if_else(
      is_missing_ppn(ppn) & !is.na(survey_suggested_ppn) & trimws(survey_suggested_ppn) != "",
      survey_suggested_ppn,
      ppn
    )
  )

# -------------------------
# 5) Build Demo suggestions
#    Match Demo StartDate to IC EndDate, then inherit IC ResponseId (canonical PPN).
# -------------------------
if (need_fix_demo) {

  demo_missing <- demo_B %>%
    dplyr::filter(is_missing_ppn(ppn)) %>%
    dplyr::transmute(
      component = "demo",
      row_id,
      StartDate_posix,
      EndDate_posix,
      match_target_time = StartDate_posix
    )

  demo_review_tbl <- demo_missing %>%
    dplyr::rowwise() %>%
    dplyr::mutate(
      .m = list(time_match_one(match_target_time, ic_B$EndDate_posix, tolerance_secs)),
      ic_row_id = .m$best_index,
      ic_best_abs_diff = .m$best_abs_diff,
      ic_n_within_tol = .m$n_within_tol,
      ic_within_tol = !is.na(ic_row_id),

      ic_StartDate_posix = if (ic_within_tol) ic_B$StartDate_posix[ic_row_id] else as.POSIXct(NA),
      ic_EndDate_posix   = if (ic_within_tol) ic_B$EndDate_posix[ic_row_id] else as.POSIXct(NA),
      ic_ppn             = if (ic_within_tol) ic_B$ppn_ic[ic_row_id] else NA_character_,

      # signed delta for audit (ic end - demo start)
      ic_time_diff_secs = if (ic_within_tol)
        as.numeric(difftime(ic_EndDate_posix, match_target_time, units = "secs"))
      else NA_real_
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      suggested_ppn = dplyr::case_when(
        ic_within_tol & !is.na(ic_ppn) & trimws(ic_ppn) != "" ~ ic_ppn,
        TRUE ~ NA_character_
      ),
      suggestion_status = dplyr::case_when(
        is.na(StartDate_posix) ~ "missing_demo_startdate",
        is.na(ic_row_id) ~ "no_ic_time_match",
        is.na(ic_ppn) | trimws(ic_ppn) == "" ~ "matched_ic_but_ic_ppn_missing",
        TRUE ~ "inherited_from_ic"
      ),
      recommend_apply = !is.na(suggested_ppn) & trimws(suggested_ppn) != ""
    ) %>%
    dplyr::select(
      component, row_id,
      StartDate_posix, EndDate_posix,

      # IC evidence
      ic_row_id,
      ic_StartDate_posix, ic_EndDate_posix,
      ic_ppn,
      ic_time_diff_secs,
      ic_n_within_tol,

      # suggestion
      suggested_ppn,
      suggestion_status,
      recommend_apply
    )
}

# -------------------------
# 6) Write ONE combined review file (even if empty)
# -------------------------
review_tbl <- dplyr::bind_rows(
  survey_review_tbl,
  demo_review_tbl
)

review_out <- review_tbl %>%
  dplyr::mutate(
    apply_fix = recommend_apply,
    ppn_new = "",
    notes = ""
  ) %>%
  dplyr::arrange(component, dplyr::desc(recommend_apply))

if (nrow(review_out) == 0) {
  review_out <- tibble::tibble(
    component = character(),
    row_id = integer(),
    StartDate_posix = as.POSIXct(character()),
    EndDate_posix = as.POSIXct(character()),
    suggested_ppn = character(),
    suggestion_status = character(),
    recommend_apply = logical(),
    apply_fix = logical(),
    ppn_new = character(),
    notes = character()
  )
}

write_review_csv(review_out, "B_missing_ppn_review.csv")

# -------------------------
# 7) Re-import edited review + apply fixes
# -------------------------
review_edited <- read_audit_csv2("B_missing_ppn_review.csv") %>%
  dplyr::mutate(
    component = as.character(component),
    row_id = suppressWarnings(as.integer(row_id)),
    apply_fix = as.logical(apply_fix),
    ppn_new = as.character(ppn_new),
    suggested_ppn = as.character(suggested_ppn),
    ppn_to_apply = dplyr::if_else(
      !is.na(ppn_new) & trimws(ppn_new) != "",
      trimws(ppn_new),
      trimws(suggested_ppn)
    )
  )

stopifnot(!any(is.na(review_edited$row_id)))

fixes_to_apply <- review_edited %>%
  dplyr::filter(apply_fix == TRUE) %>%
  dplyr::filter(!is.na(ppn_to_apply) & trimws(ppn_to_apply) != "") %>%
  dplyr::select(component, row_id, ppn_to_apply, suggestion_status, recommend_apply, notes)

# Always define working outputs
survey_working <- survey_B
demo_working   <- demo_B

if (nrow(fixes_to_apply) > 0) {

  # ---- (1) apply Survey fixes ----
  survey_fixes <- fixes_to_apply %>% dplyr::filter(component == "survey")
  if (nrow(survey_fixes) > 0) {
    survey_working <- survey_B %>%
      dplyr::left_join(survey_fixes %>% dplyr::select(row_id, ppn_to_apply), by = "row_id") %>%
      dplyr::mutate(
        ppn_before = ppn,
        ppn = dplyr::if_else(is_missing_ppn(ppn) & !is.na(ppn_to_apply) & trimws(ppn_to_apply) != "", ppn_to_apply, ppn)
      ) %>%
      dplyr::select(-ppn_to_apply)

    survey_receipt <- survey_B %>%
      dplyr::mutate(ppn_before = ppn) %>%
      dplyr::left_join(survey_fixes, by = "row_id") %>%
      dplyr::left_join(
        survey_working %>% dplyr::select(row_id, ppn_after = ppn),
        by = "row_id"
      ) %>%
      dplyr::filter(is_missing_ppn(ppn_before) & !is.na(ppn_to_apply) & trimws(ppn_to_apply) != "") %>%
      dplyr::transmute(component = "survey", row_id, ppn_before, ppn_after, ppn_to_apply, suggestion_status, recommend_apply, notes)

    write_receipt_csv(survey_receipt, "B_missing_ppn_receipt_survey.csv")
  }

  # ---- (2) apply Demo fixes ----
  demo_fixes <- fixes_to_apply %>% dplyr::filter(component == "demo")
  if (nrow(demo_fixes) > 0) {
    demo_working <- demo_B %>%
      dplyr::left_join(demo_fixes %>% dplyr::select(row_id, ppn_to_apply), by = "row_id") %>%
      dplyr::mutate(
        ppn_before = ppn,
        ppn = dplyr::if_else(is_missing_ppn(ppn) & !is.na(ppn_to_apply) & trimws(ppn_to_apply) != "", ppn_to_apply, ppn)
      ) %>%
      dplyr::select(-ppn_to_apply)

    demo_receipt <- demo_B %>%
      dplyr::mutate(ppn_before = ppn) %>%
      dplyr::left_join(demo_fixes, by = "row_id") %>%
      dplyr::left_join(
        demo_working %>% dplyr::select(row_id, ppn_after = ppn),
        by = "row_id"
      ) %>%
      dplyr::filter(is_missing_ppn(ppn_before) & !is.na(ppn_to_apply) & trimws(ppn_to_apply) != "") %>%
      dplyr::transmute(component = "demo", row_id, ppn_before, ppn_after, ppn_to_apply, suggestion_status, recommend_apply, notes)

    write_receipt_csv(demo_receipt, "B_missing_ppn_receipt_demo.csv")
  }
}

# -------------------------
# 8) Final objects exported by this step (for downstream sections)
# -------------------------
# Use these moving forward:
#   - survey_working
#   - demo_working
#
# Keep these for traceability:
#   - review_out (written)
#   - review_edited
#   - fixes_to_apply
###############################################################################

```



```{r}
###############################################################################
# AUDIT STEP C (Wave 0): resolve shared ppn within downstream dataset
#
# Goal
# Detect cases where the same ppn appears more than once WITHIN demo or
# WITHIN survey. This usually indicates that a participant shared their
# survey link after completing IC.
#
# Because Wave 0 lacks the financial compensation anchor used in Wave 1,
# we attempt to recover the most plausible continuation of the IC case
# using timing continuity.
#
# Strategy
#   C1. Detect duplicated ppns within demo or survey.
#   C2. Attach IC anchor timing evidence.
#   C3. Rank rows by temporal plausibility.
#   C4. Auto-mark the best candidate as keep=TRUE.
#   C5. Auto-mark all other rows as keep=FALSE and flag_review=TRUE.
#   C6. Write one review file for manual inspection.
#   C7. Read edited file back in and apply keep/drop decisions.
#   C8. Write one receipt showing rows actually removed.
#
# Outputs
#   - C_shared_ppn_review.csv
#   - receipts/C_shared_ppn_receipt.csv
###############################################################################

# -------------------------
# C0) Choose working objects
# -------------------------

demo_C <- if (exists("demo_working")) demo_working else demo_raw_wave0
survey_C <- if (exists("survey_working")) survey_working else survey_wave0

demo_C <- demo_C %>%
  mutate(
    row_id = if ("row_id" %in% names(.)) row_id else dplyr::row_number(),
    ppn = as.character(ppn),
    StartDate_posix = if ("StartDate_posix" %in% names(.)) StartDate_posix else parse_dt_any(StartDate),
    EndDate_posix   = if ("EndDate_posix" %in% names(.)) EndDate_posix else parse_dt_any(EndDate)
  )

survey_C <- survey_C %>%
  mutate(
    row_id = if ("row_id" %in% names(.)) row_id else dplyr::row_number(),
    ppn = as.character(ppn),
    StartDate_posix = if ("StartDate_posix" %in% names(.)) StartDate_posix else parse_dt_any(StartDate),
    EndDate_posix   = if ("EndDate_posix" %in% names(.)) EndDate_posix else parse_dt_any(EndDate)
  )

# IC anchor
ic_C <- ic_combined %>%
  mutate(
    ic_ppn = as.character(ResponseId),
    ic_name = dplyr::coalesce(
      if ("Name_title" %in% names(.)) as.character(Name_title) else NA_character_,
      if ("Name" %in% names(.)) as.character(Name) else NA_character_,
      if ("Name_clean" %in% names(.)) as.character(Name_clean) else NA_character_
    ),
    ic_EndDate_posix = if ("EndDate_posix" %in% names(.)) EndDate_posix else parse_dt_any(EndDate)
  ) %>%
  select(ic_ppn, ic_name, ic_EndDate_posix)

# -------------------------
# C1) Detect duplicated ppn within dataset
# -------------------------

downstream_C <- bind_rows(
  demo_C %>% mutate(component = "demo"),
  survey_C %>% mutate(component = "survey")
)

shared_ppn_summary <- downstream_C %>%
  filter(!is_missing_ppn(ppn)) %>%
  group_by(component, ppn) %>%
  summarise(n_rows = n(), .groups = "drop") %>%
  filter(n_rows > 1)

shared_keys <- shared_ppn_summary %>%
  mutate(shared_key = paste(component, ppn, sep = "|||")) %>%
  pull(shared_key)

candidate_rows <- downstream_C %>%
  mutate(shared_key = paste(component, ppn, sep = "|||")) %>%
  filter(shared_key %in% shared_keys) %>%
  left_join(ic_C, by = c("ppn" = "ic_ppn")) %>%
  arrange(component, ppn, row_id)

# -------------------------
# C2) Attach timing evidence
# -------------------------

candidate_rows_scored <- candidate_rows %>%
  mutate(
    ic_to_start_diff_secs =
      abs(as.numeric(difftime(StartDate_posix, ic_EndDate_posix, units = "secs")))
  ) %>%
  group_by(component, ppn) %>%
  mutate(
    evidence_score = ic_to_start_diff_secs,
    rank_within_shared_ppn =
      rank(evidence_score, ties.method = "first", na.last = "keep"),
    n_rows_with_same_ppn = n()
  ) %>%
  ungroup()

# -------------------------
# C3) WRITE review file
# -------------------------

review_out <- candidate_rows_scored %>%
  mutate(
    keep = rank_within_shared_ppn == 1,
    flag_review = rank_within_shared_ppn != 1,
    keep_reason = case_when(
      rank_within_shared_ppn == 1 ~ "best_timing_match_to_ic",
      TRUE ~ "shared_ppn_secondary_row"
    )
  ) %>%
  transmute(
    component,
    row_id,
    ppn_current = ppn,
    StartDate_posix,
    EndDate_posix,
    ic_name,
    ic_EndDate_posix,
    ic_to_start_diff_secs,
    evidence_score,
    rank_within_shared_ppn,
    n_rows_with_same_ppn,

    keep,
    flag_review,
    keep_reason,
    notes = ""
  ) %>%
  arrange(component, ppn_current, rank_within_shared_ppn)

write_review_csv(review_out, "C_shared_ppn_review.csv")

# -------------------------
# C4) READ edited review file
# -------------------------

review_edited <- read_audit_csv2("C_shared_ppn_review.csv") %>%
  mutate(
    component = as.character(component),
    row_id = as.integer(row_id),
    keep = as.logical(keep),
    flag_review = as.logical(flag_review),
    keep_reason = as.character(keep_reason),
    notes = as.character(notes)
  )

stopifnot(!any(is.na(review_edited$row_id)))

rows_to_drop <- review_edited %>%
  filter(keep == FALSE)

# -------------------------
# C5) APPLY decisions
# -------------------------

demo_working_C <- demo_C
survey_working_C <- survey_C

demo_drop <- rows_to_drop %>% filter(component == "demo")

if (nrow(demo_drop) > 0) {
  demo_working_C <- demo_C %>%
    anti_join(demo_drop %>% select(row_id), by = "row_id")
}

survey_drop <- rows_to_drop %>% filter(component == "survey")

if (nrow(survey_drop) > 0) {
  survey_working_C <- survey_C %>%
    anti_join(survey_drop %>% select(row_id), by = "row_id")
}

demo_working <- demo_working_C
survey_working <- survey_working_C

# -------------------------
# C6) WRITE receipt
# -------------------------

receipt_out <- rows_to_drop %>%
  transmute(
    component,
    row_id,
    keep = FALSE,
    flag_review,
    keep_reason,
    notes
  ) %>%
  arrange(component, row_id)

write_receipt_csv(receipt_out, "C_shared_ppn_receipt.csv")

###############################################################################
# Final objects exported by this step
#   demo_working
#   survey_working
###############################################################################



```



```{r}
###############################################################################
# AUDIT STEP D (Wave 0): IC restarters (same name, multiple IC ResponseIds)
###############################################################################
# Purpose:
#   Detect IC "restarts" (same cleaned name submitted IC more than once) and use
#   downstream evidence to select ONE canonical IC ResponseId per identity.
#
# Key idea:
#   IC ResponseId is treated as the wave canonical participant identifier (ppn).
#   When a person restarts IC, we must choose a single canonical ResponseId so
#   that all downstream linkage is stable.
#
# Inputs expected to exist by this stage:
#   - ic_ledger_working  (IC after Audit Step A test removals)
#   - demo_working2      (after missing ppn repairs and shared ppn repairs)
#   - survey_working2
#
# Output strategy (audit_dir):
#   - D_identity_ledger.csv        (read only, one row per identity)
#   - D_identity_review_queue.csv  (editable, only restart cases needing review)
#
# Receipt output (receipts_dir if available, otherwise audit_dir):
#   - D_identity_review_applied.csv  (only override and drop decisions)
#
# Review decisions (editable in D_identity_review_queue.csv):
#   - decision: accept / override / drop
#   - canonical_responseid_final: fill only if decision == override
#   - drop_reason, notes
###############################################################################

# ---------------------------------------------------------------------------
# 0) Pick working datasets with safe fallbacks
# ---------------------------------------------------------------------------
stopifnot(exists("ic_ledger_working"))

demo_D   <- if (exists("demo_working2")) demo_working2 else if (exists("demo_working")) demo_working else if (exists("demo_B")) demo_B else demo_raw_wave0
survey_D <- if (exists("survey_working2")) survey_working2 else if (exists("survey_working")) survey_working else if (exists("survey_B")) survey_B else survey_wave0

# Normalize ppn vectors for presence checks
demo_ppn_vec   <- unique(trimws(as.character(demo_D$ppn)))
survey_ppn_vec <- unique(trimws(as.character(survey_D$ppn)))

# ---------------------------------------------------------------------------
# 1) Build ic_presence table: one row per IC ResponseId candidate
# ---------------------------------------------------------------------------
# Expected columns in ic_ledger_working from Step A:
#   Name_clean, Name_title, ResponseId, flag_test_by_date, flag_test_by_name, StartDate_posix
stopifnot(all(c("Name_clean", "Name_title", "ResponseId") %in% names(ic_ledger_working)))

# If StartDate_posix is missing for some reason, reconstruct it
if (!("StartDate_posix" %in% names(ic_ledger_working))) {
  ic_ledger_working <- ic_ledger_working %>%
    dplyr::mutate(StartDate_posix = parse_dt_any(StartDate))
}

ic_presence <- ic_ledger_working %>%
  dplyr::distinct(Name_clean, Name_title, ResponseId, flag_test_by_date, flag_test_by_name, StartDate_posix) %>%
  dplyr::mutate(
    ResponseId = trimws(as.character(ResponseId)),
    in_demo    = ResponseId %in% demo_ppn_vec,
    in_survey  = ResponseId %in% survey_ppn_vec
  ) %>%
  dplyr::mutate(
    # Wave 0 downstream scoring:
    # demo + survey are the only evidence available.
    downstream_score =
      (in_demo   * 3L) +   # treat demo as strong evidence
      (in_survey * 3L)     # treat survey as strong evidence
  )

# ---------------------------------------------------------------------------
# 2) Rank candidate ResponseIds within each Name_clean cluster
# ---------------------------------------------------------------------------
# Tie-break: higher downstream_score wins; then earlier IC start (conservative),
# then stable ordering by ResponseId.
ic_ranked <- ic_presence %>%
  dplyr::group_by(Name_clean) %>%
  dplyr::arrange(
    dplyr::desc(downstream_score),
    StartDate_posix,
    ResponseId,
    .by_group = TRUE
  ) %>%
  dplyr::mutate(
    n_ic_rows_for_name = dplyr::n(),
    canonical_responseid_auto = dplyr::first(ResponseId),
    canonical_score_auto      = downstream_score[match(canonical_responseid_auto, ResponseId)],
    is_canonical_auto = ResponseId == canonical_responseid_auto
  ) %>%
  dplyr::ungroup()

# ---------------------------------------------------------------------------
# 3) One row per identity ledger: full record (read only)
# ---------------------------------------------------------------------------
identity_ledger <- ic_ranked %>%
  dplyr::group_by(Name_clean) %>%
  dplyr::summarise(
    Name_title = dplyr::first(stats::na.omit(Name_title)),
    n_ic_rows  = dplyr::first(n_ic_rows_for_name),

    canonical_responseid_auto = dplyr::first(canonical_responseid_auto),
    canonical_score_auto      = dplyr::first(canonical_score_auto),

    max_score = suppressWarnings(max(downstream_score, na.rm = TRUE)),
    n_at_max  = suppressWarnings(sum(downstream_score == max_score, na.rm = TRUE)),

    ic_start_min = suppressWarnings(min(StartDate_posix, na.rm = TRUE)),
    ic_start_max = suppressWarnings(max(StartDate_posix, na.rm = TRUE)),

    any_test_flags = any(flag_test_by_date | flag_test_by_name, na.rm = TRUE),

    n_candidates_in_demo   = sum(in_demo, na.rm = TRUE),
    n_candidates_in_survey = sum(in_survey, na.rm = TRUE),

    all_responseids = paste(ResponseId, collapse = " | "),
    scores_by_responseid = paste0(ResponseId, ":", downstream_score) %>% paste(collapse = " | "),
    demo_by_responseid    = paste0(ResponseId, ":", as.integer(in_demo))   %>% paste(collapse = " | "),
    survey_by_responseid  = paste0(ResponseId, ":", as.integer(in_survey)) %>% paste(collapse = " | "),

    reason_summary = dplyr::case_when(
      n_ic_rows == 1 ~ "single_ic_row",
      n_ic_rows > 1 & any(downstream_score > 0, na.rm = TRUE) ~ "multiple_ic_rows_pick_best_downstream",
      n_ic_rows > 1 ~ "multiple_ic_rows_no_downstream_signal",
      TRUE ~ "unknown"
    ),

    .groups = "drop"
  ) %>%
  dplyr::mutate(
    ic_start_min = dplyr::if_else(is.infinite(ic_start_min), as.POSIXct(NA, tz = "UTC"), ic_start_min),
    ic_start_max = dplyr::if_else(is.infinite(ic_start_max), as.POSIXct(NA, tz = "UTC"), ic_start_max),

    flag_restart    = n_ic_rows > 1,
    flag_tie_at_max = flag_restart & n_at_max > 1,
    flag_low_signal = canonical_score_auto == 0,

    # Wave 0 auto_ok threshold:
    # max score possible is 6 (demo 3 + survey 3). Treat 6 as "high confidence".
    auto_ok = canonical_score_auto >= 6 & !flag_tie_at_max & !any_test_flags,

    review_priority = dplyr::case_when(
      any_test_flags ~ "review_high_test",
      flag_tie_at_max ~ "review_high_tie",
      canonical_score_auto == 0 ~ "review_high_no_signal",
      canonical_score_auto < 6 ~ "review_medium_low_score",
      TRUE ~ "auto_ok"
    )
  ) %>%
  dplyr::arrange(review_priority, Name_title)

write_audit_csv(identity_ledger, "D_identity_ledger.csv")

# ---------------------------------------------------------------------------
# 4) Restart-only review queue: editable, only cases needing review
# ---------------------------------------------------------------------------
identity_review_queue_template <- identity_ledger %>%
  dplyr::filter(
    flag_restart == TRUE,
    review_priority != "auto_ok"
  ) %>%
  dplyr::transmute(
    Name_title,
    Name_clean,

    n_ic_rows,
    canonical_responseid_auto,
    canonical_score_auto,
    max_score,
    n_at_max,

    ic_start_min,
    ic_start_max,

    n_candidates_in_demo,
    n_candidates_in_survey,

    all_responseids,
    scores_by_responseid,
    demo_by_responseid,
    survey_by_responseid,

    any_test_flags,
    reason_summary,
    flag_tie_at_max,
    flag_low_signal,
    review_priority,

    decision = "accept",
    canonical_responseid_final = "",
    drop_reason = "",
    notes = ""
  )

write_review_csv(identity_review_queue_template, "D_identity_review_queue.csv")

# ---------------------------------------------------------------------------
# 5) Re-import decisions and apply to produce identity_roster
# ---------------------------------------------------------------------------
identity_review_queue <- read_audit_csv2("D_identity_review_queue.csv")

# Safety checks: reviewer should not reorder or delete rows
stopifnot(nrow(identity_review_queue) == nrow(identity_review_queue_template))
stopifnot(all(identity_review_queue$Name_clean == identity_review_queue_template$Name_clean))
stopifnot(all(identity_review_queue$canonical_responseid_auto == identity_review_queue_template$canonical_responseid_auto))

identity_roster <- identity_ledger %>%
  dplyr::left_join(
    identity_review_queue %>%
      dplyr::select(Name_clean, decision, canonical_responseid_final, drop_reason, notes),
    by = "Name_clean"
  ) %>%
  dplyr::mutate(
    decision = dplyr::if_else(is.na(decision) | decision == "", "accept", as.character(decision)),

    canonical_responseid_final = dplyr::case_when(
      decision == "override" &
        !is.na(canonical_responseid_final) &
        canonical_responseid_final != "" ~ as.character(canonical_responseid_final),

      decision == "accept" ~ as.character(canonical_responseid_auto),

      decision == "drop" ~ NA_character_,

      TRUE ~ as.character(canonical_responseid_auto)
    ),

    override_is_valid = dplyr::if_else(
      decision == "override" &
        !is.na(canonical_responseid_final) &
        canonical_responseid_final != "",
      stringr::str_detect(all_responseids, stringr::fixed(canonical_responseid_final)),
      TRUE
    )
  )

stopifnot(all(identity_roster$override_is_valid))

# Receipt: only rows where reviewer changed something
identity_review_applied <- identity_roster %>%
  dplyr::filter(decision %in% c("override", "drop")) %>%
  dplyr::select(
    Name_title, Name_clean,
    decision,
    canonical_responseid_auto, canonical_score_auto,
    canonical_responseid_final,
    drop_reason, notes,
    n_ic_rows, all_responseids, scores_by_responseid,
    review_priority, flag_tie_at_max, flag_low_signal, any_test_flags
  )

if (exists("write_receipt_csv")) {
  write_receipt_csv(identity_review_applied, "D_identity_review_applied.csv")
} else {
  write_audit_csv(identity_review_applied, "D_identity_review_applied.csv")
}

###############################################################################
# End result:
#   identity_roster contains canonical_responseid_final (IC ResponseId as ppn)
#   Use this in later steps to enforce canonical ppns across demo/survey.
###############################################################################

```



```{r}
###############################################################################
# AUDIT STEP E (Wave 0): incompletes (canonical_score_auto < 6) + optional exclusion
###############################################################################
# Output strategy:
#   - One audit file (editable, in audit_dir):
#       E_incompletes_review.csv
#   - One receipt file (read-only, in receipts_dir):
#       E_incompletes_receipt.csv
#
# Rules:
#   - The audit file contains ONLY incompletes (canonical_score_auto < 6)
#   - IC-only identities are default keep = FALSE
#   - Any identity with any downstream rows defaults keep = TRUE
#   - If any are dropped, rebuild identity_ledger and identity_roster in place
###############################################################################

stopifnot(exists("identity_ledger"))
stopifnot(exists("identity_roster"))
stopifnot(exists("ic_ledger_working"))
stopifnot(exists("write_audit_csv"))
stopifnot(exists("write_receipt_csv"))

# ---------------------------------------------------------------------------
# 0) Choose most current working datasets
# ---------------------------------------------------------------------------
demo_E   <- if (exists("demo_working2")) demo_working2 else if (exists("demo_working")) demo_working else if (exists("demo_B")) demo_B else demo_raw_wave0
survey_E <- if (exists("survey_working2")) survey_working2 else if (exists("survey_working")) survey_working else if (exists("survey_B")) survey_B else survey_wave0

demo_E   <- demo_E   %>% dplyr::mutate(ppn = trimws(as.character(ppn)))
survey_E <- survey_E %>% dplyr::mutate(ppn = trimws(as.character(ppn)))

# ---------------------------------------------------------------------------
# 1) Stage summary helper (rows, progress, finished, status) by ppn
# ---------------------------------------------------------------------------
summarise_stage_by_ppn <- function(df, stage_prefix) {
  if (is.null(df) || nrow(df) == 0) {
    return(tibble::tibble(
      canonical_ppn = character(),
      !!paste0(stage_prefix, "_n_rows") := integer(),
      !!paste0(stage_prefix, "_progress_max") := numeric(),
      !!paste0(stage_prefix, "_finished_any") := logical(),
      !!paste0(stage_prefix, "_status") := character()
    ))
  }

  df <- df %>% dplyr::mutate(ppn = trimws(as.character(ppn)))

  has_progress <- "Progress" %in% names(df)
  has_finished <- "Finished" %in% names(df)

  df %>%
    dplyr::filter(!is.na(ppn) & ppn != "") %>%
    dplyr::group_by(ppn) %>%
    dplyr::summarise(
      n_rows = dplyr::n(),
      progress_max = if (has_progress) suppressWarnings(max(as.numeric(Progress), na.rm = TRUE)) else NA_real_,
      finished_any = if (has_finished) any(suppressWarnings(as.numeric(Finished)) == 1, na.rm = TRUE) else NA,
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      stage_status = dplyr::case_when(
        n_rows == 0 ~ "missing",
        !is.na(finished_any) & finished_any ~ "complete",
        !is.na(progress_max) & progress_max >= 100 ~ "complete",
        TRUE ~ "partial"
      )
    ) %>%
    dplyr::transmute(
      canonical_ppn = ppn,
      !!paste0(stage_prefix, "_n_rows") := as.integer(n_rows),
      !!paste0(stage_prefix, "_progress_max") := as.numeric(progress_max),
      !!paste0(stage_prefix, "_finished_any") := as.logical(finished_any),
      !!paste0(stage_prefix, "_status") := as.character(stage_status)
    )
}

# IC stage: ResponseId is canonical ppn
ic_stage <- ic_ledger_working %>%
  dplyr::mutate(ResponseId = trimws(as.character(ResponseId))) %>%
  dplyr::filter(!is.na(ResponseId) & ResponseId != "") %>%
  dplyr::group_by(ResponseId) %>%
  dplyr::summarise(ic_n_rows = dplyr::n(), .groups = "drop") %>%
  dplyr::transmute(
    canonical_ppn = ResponseId,
    ic_n_rows = as.integer(ic_n_rows),
    ic_progress_max = as.numeric(NA),
    ic_finished_any = as.logical(NA),
    ic_status = "present"
  )

demo_stage   <- summarise_stage_by_ppn(demo_E,   "demo")
survey_stage <- summarise_stage_by_ppn(survey_E, "survey")

stage_summary <- ic_stage %>%
  dplyr::full_join(demo_stage,   by = "canonical_ppn") %>%
  dplyr::full_join(survey_stage, by = "canonical_ppn") %>%
  dplyr::mutate(
    ic_n_rows      = ifelse(is.na(ic_n_rows), 0L, ic_n_rows),
    demo_n_rows    = ifelse(is.na(demo_n_rows), 0L, demo_n_rows),
    survey_n_rows  = ifelse(is.na(survey_n_rows), 0L, survey_n_rows),

    ic_status      = ifelse(is.na(ic_status) & ic_n_rows > 0, "present", ic_status),
    demo_status    = ifelse(is.na(demo_status)   & demo_n_rows   > 0, "partial", demo_status),
    survey_status  = ifelse(is.na(survey_status) & survey_n_rows > 0, "partial", survey_status),

    ic_status      = ifelse(is.na(ic_status), "missing", ic_status),
    demo_status    = ifelse(is.na(demo_status), "missing", demo_status),
    survey_status  = ifelse(is.na(survey_status), "missing", survey_status)
  )

# ---------------------------------------------------------------------------
# 2) Build ONE review table (ONLY canonical_score_auto < 6)
#    Use roster canonical_responseid_final as canonical ppn per identity.
# ---------------------------------------------------------------------------
roster_ppn <- identity_roster %>%
  dplyr::mutate(
    decision = ifelse(is.na(decision) | decision == "", "accept", as.character(decision)),
    canonical_responseid_final = trimws(as.character(canonical_responseid_final))
  ) %>%
  dplyr::filter(decision != "drop") %>%
  dplyr::transmute(
    Name_clean,
    canonical_ppn = canonical_responseid_final
  ) %>%
  dplyr::filter(!is.na(canonical_ppn) & canonical_ppn != "") %>%
  dplyr::distinct(Name_clean, .keep_all = TRUE)

E_review_template <- identity_ledger %>%
  dplyr::left_join(roster_ppn, by = "Name_clean") %>%
  dplyr::filter(!is.na(canonical_ppn) & canonical_ppn != "") %>%
  dplyr::filter(canonical_score_auto < 6) %>%
  dplyr::left_join(stage_summary, by = "canonical_ppn") %>%
  dplyr::mutate(
    downstream_any_rows = (demo_n_rows + survey_n_rows) > 0L,
    keep = downstream_any_rows,
    drop_reason = dplyr::case_when(
      downstream_any_rows == FALSE ~ "ic_only_no_usable_downstream_data",
      TRUE ~ ""
    ),
    notes = ""
  ) %>%
  dplyr::transmute(
    Name_title,
    Name_clean,
    canonical_ppn,

    canonical_score_auto,
    max_score,
    review_priority,

    # Wave 0 downstream evidence columns from Step D
    n_candidates_in_demo,
    n_candidates_in_survey,

    scores_by_responseid,
    demo_by_responseid,
    survey_by_responseid,

    ic_n_rows, ic_status,
    demo_n_rows, demo_progress_max, demo_finished_any, demo_status,
    survey_n_rows, survey_progress_max, survey_finished_any, survey_status,

    downstream_any_rows,

    keep, drop_reason, notes
  ) %>%
  dplyr::arrange(canonical_score_auto, Name_title)

write_review_csv(E_review_template, "E_incompletes_review.csv")

# ---------------------------------------------------------------------------
# 3) Read decisions and compute drops
# ---------------------------------------------------------------------------
E_review <- read_audit_csv2("E_incompletes_review.csv") %>%
  dplyr::mutate(
    Name_clean = as.character(Name_clean),
    canonical_ppn = trimws(as.character(canonical_ppn)),
    keep = as.logical(keep),
    drop_reason = as.character(drop_reason),
    notes = as.character(notes)
  )

stopifnot(nrow(E_review) == nrow(E_review_template))
stopifnot(all(E_review$Name_clean == E_review_template$Name_clean))
stopifnot(all(E_review$canonical_ppn == E_review_template$canonical_ppn))

drop_ppns_E <- E_review %>%
  dplyr::filter(keep == FALSE) %>%
  dplyr::pull(canonical_ppn) %>%
  unique()

drop_ppns_E <- drop_ppns_E[!is.na(drop_ppns_E) & drop_ppns_E != ""]

# ---------------------------------------------------------------------------
# 4) Apply drops only if needed, then rebuild identity_ledger and identity_roster
# ---------------------------------------------------------------------------
n_drop <- length(drop_ppns_E)

if (n_drop > 0) {

  # Filter working datasets
  ic_ledger_working <- ic_ledger_working %>%
    dplyr::mutate(ResponseId = trimws(as.character(ResponseId))) %>%
    dplyr::filter(!(ResponseId %in% drop_ppns_E))

  demo_E   <- demo_E   %>% dplyr::filter(!(ppn %in% drop_ppns_E))
  survey_E <- survey_E %>% dplyr::filter(!(ppn %in% drop_ppns_E))

  # Recompute presence vectors
  demo_ppn_vec   <- unique(demo_E$ppn)
  survey_ppn_vec <- unique(survey_E$ppn)

  # Rebuild ic_presence using Wave 0 scoring
  ic_presence <- ic_ledger_working %>%
    dplyr::distinct(Name_clean, Name_title, ResponseId, flag_test_by_date, flag_test_by_name, StartDate_posix) %>%
    dplyr::mutate(
      ResponseId = trimws(as.character(ResponseId)),
      in_demo    = ResponseId %in% demo_ppn_vec,
      in_survey  = ResponseId %in% survey_ppn_vec
    ) %>%
    dplyr::mutate(
      downstream_score =
        (in_demo   * 3L) +
        (in_survey * 3L)
    )

  ic_ranked <- ic_presence %>%
    dplyr::group_by(Name_clean) %>%
    dplyr::arrange(dplyr::desc(downstream_score), StartDate_posix, ResponseId, .by_group = TRUE) %>%
    dplyr::mutate(
      n_ic_rows_for_name = dplyr::n(),
      canonical_responseid_auto = dplyr::first(ResponseId),
      canonical_score_auto = downstream_score[match(canonical_responseid_auto, ResponseId)],
      is_canonical_auto = ResponseId == canonical_responseid_auto
    ) %>%
    dplyr::ungroup()

  identity_ledger <- ic_ranked %>%
    dplyr::group_by(Name_clean) %>%
    dplyr::summarise(
      Name_title = dplyr::first(stats::na.omit(Name_title)),
      n_ic_rows  = dplyr::first(n_ic_rows_for_name),

      canonical_responseid_auto = dplyr::first(canonical_responseid_auto),
      canonical_score_auto      = dplyr::first(canonical_score_auto),

      max_score = suppressWarnings(max(downstream_score, na.rm = TRUE)),
      n_at_max  = suppressWarnings(sum(downstream_score == max_score, na.rm = TRUE)),

      ic_start_min = suppressWarnings(min(StartDate_posix, na.rm = TRUE)),
      ic_start_max = suppressWarnings(max(StartDate_posix, na.rm = TRUE)),

      any_test_flags = any(flag_test_by_date | flag_test_by_name, na.rm = TRUE),

      n_candidates_in_demo   = sum(in_demo, na.rm = TRUE),
      n_candidates_in_survey = sum(in_survey, na.rm = TRUE),

      all_responseids = paste(ResponseId, collapse = " | "),
      scores_by_responseid = paste0(ResponseId, ":", downstream_score) %>% paste(collapse = " | "),
      demo_by_responseid    = paste0(ResponseId, ":", as.integer(in_demo))   %>% paste(collapse = " | "),
      survey_by_responseid  = paste0(ResponseId, ":", as.integer(in_survey)) %>% paste(collapse = " | "),

      reason_summary = dplyr::case_when(
        n_ic_rows == 1 ~ "single_ic_row",
        n_ic_rows > 1 & any(downstream_score > 0, na.rm = TRUE) ~ "multiple_ic_rows_pick_best_downstream",
        n_ic_rows > 1 ~ "multiple_ic_rows_no_downstream_signal",
        TRUE ~ "unknown"
      ),

      .groups = "drop"
    ) %>%
    dplyr::mutate(
      ic_start_min = dplyr::if_else(is.infinite(ic_start_min), as.POSIXct(NA, tz = "UTC"), ic_start_min),
      ic_start_max = dplyr::if_else(is.infinite(ic_start_max), as.POSIXct(NA, tz = "UTC"), ic_start_max),

      flag_restart    = n_ic_rows > 1,
      flag_tie_at_max = flag_restart & n_at_max > 1,
      flag_low_signal = canonical_score_auto == 0,

      auto_ok = canonical_score_auto >= 6 & !flag_tie_at_max & !any_test_flags,

      review_priority = dplyr::case_when(
        any_test_flags ~ "review_high_test",
        flag_tie_at_max ~ "review_high_tie",
        canonical_score_auto == 0 ~ "review_high_no_signal",
        canonical_score_auto < 6 ~ "review_medium_low_score",
        TRUE ~ "auto_ok"
      )
    ) %>%
    dplyr::arrange(review_priority, Name_title)

  # Preserve any existing manual restarter decisions already in identity_roster
  manual_decisions <- identity_roster %>%
    dplyr::transmute(
      Name_clean,
      decision_manual = as.character(decision),
      canonical_responseid_final_manual = as.character(canonical_responseid_final),
      drop_reason_manual = as.character(drop_reason),
      notes_manual = as.character(notes)
    )

  identity_roster <- identity_ledger %>%
    dplyr::left_join(manual_decisions, by = "Name_clean") %>%
    dplyr::mutate(
      decision = ifelse(!is.na(decision_manual) & decision_manual != "", decision_manual, "accept"),
      canonical_responseid_final = dplyr::case_when(
        decision == "override" &
          !is.na(canonical_responseid_final_manual) &
          canonical_responseid_final_manual != "" ~ canonical_responseid_final_manual,
        decision == "drop" ~ NA_character_,
        TRUE ~ canonical_responseid_auto
      ),
      drop_reason = ifelse(!is.na(drop_reason_manual) & drop_reason_manual != "", drop_reason_manual, ""),
      notes = ifelse(!is.na(notes_manual) & notes_manual != "", notes_manual, "")
    ) %>%
    dplyr::select(-dplyr::ends_with("_manual"))

  identity_roster <- identity_roster %>%
    dplyr::mutate(
      override_is_valid = dplyr::if_else(
        decision == "override" & !is.na(canonical_responseid_final) & canonical_responseid_final != "",
        stringr::str_detect(all_responseids, stringr::fixed(canonical_responseid_final)),
        TRUE
      )
    )
  stopifnot(all(identity_roster$override_is_valid))
}

# ---------------------------------------------------------------------------
# 5) Single receipt (always) -> receipts_dir
# ---------------------------------------------------------------------------
E_receipt <- tibble::tibble(
  n_incompletes_in_review = nrow(E_review_template),
  n_marked_drop = n_drop,
  n_marked_keep = nrow(E_review_template) - n_drop
)

write_receipt_csv(E_receipt, "E_incompletes_receipt.csv")

```



```{r}
###############################################################################
# AUDIT STEP F (Wave 0): LINKAGE (Name Matching) AND gutsID PROPAGATION
###############################################################################
# Goal:
#   Attach the consortium identifier (gutsID) to each participant and propagate
#   it across Wave 0 components.
#
# Why IC is the anchor:
#   Linkage maps participant names -> gutsID.
#   IC is the only place with participant names tied to the canonical ResponseId
#   (which functions as ppn downstream).
#
# Strategy:
#   1) Standardize names in both linkage and IC into a stable join key
#   2) Apply optional private IC name corrections before linkage matching
#   3) Join IC -> linkage to assign gutsID
#   4) Build a ppn -> gutsID map from IC ResponseId
#   5) Propagate gutsID to demo / survey using ppn
#
# Outputs:
#   - One audit review file:
#       F_linkage_review.csv
#   - One receipt file (written to receipts_dir):
#       F_linkage_receipt.csv
###############################################################################

stopifnot(exists("linkage"))
stopifnot(exists("clean_name"))
stopifnot(exists("fix_tussenvoegsels"))

# ------------------------------------------------------------
# 0) Select the most current working datasets
# ------------------------------------------------------------
ic_src <- if (exists("ic_ledger_working")) {
  ic_ledger_working
} else if (exists("ic_combined_working")) {
  ic_combined_working
} else {
  ic_combined
}
stopifnot(!is.null(ic_src))

demo_src <- if (exists("demo_working2")) {
  demo_working2
} else if (exists("demo_working")) {
  demo_working
} else {
  demo_raw_wave0
}

survey_src <- if (exists("survey_working2")) {
  survey_working2
} else if (exists("survey_working")) {
  survey_working
} else {
  survey_wave0
}

demo_src <- demo_src %>%
  dplyr::mutate(ppn = trimws(as.character(ppn)))

survey_src <- survey_src %>%
  dplyr::mutate(ppn = trimws(as.character(ppn)))

# ------------------------------------------------------------
# 1) Stable title key helper
# ------------------------------------------------------------
make_title_key <- function(x) {
  x %>%
    as.character() %>%
    stringr::str_squish() %>%
    stringi::stri_trans_nfc() %>%
    stringr::str_to_title() %>%
    fix_tussenvoegsels()
}

# ------------------------------------------------------------
# 2) Build clean IC name field and apply optional private corrections
# ------------------------------------------------------------
# Participant-specific name corrections must not be hard-coded in a public
# repository. If needed, define linkage_name_patches in the private config as
# a named character vector: names = survey-side names; values = corrected names.

if (!exists("linkage_name_patches", inherits = FALSE)) {
  linkage_name_patches <- character(0)
}

ic_fixed <- ic_src %>%
  dplyr::mutate(
    ResponseId = trimws(as.character(ResponseId)),
    Name_title = dplyr::coalesce(
      if ("Name_title" %in% names(.)) as.character(Name_title) else NA_character_,
      if ("Name" %in% names(.)) as.character(Name) else NA_character_,
      if ("ic_name" %in% names(.)) as.character(ic_name) else NA_character_
    ),
    Name_title = make_title_key(Name_title),
    Name_title_original = Name_title,
    Name_title = ifelse(
      Name_title %in% names(linkage_name_patches),
      unname(linkage_name_patches[Name_title]),
      Name_title
    ),
    Name_title = make_title_key(Name_title),
    Name_title_key = clean_name(Name_title),
    hardcode_applied = Name_title != Name_title_original
  )

# ------------------------------------------------------------
# 3) Build stable linkage key
# ------------------------------------------------------------
linkage_name_col <- dplyr::case_when(
  "Name" %in% names(linkage) ~ "Name",
  "name" %in% names(linkage) ~ "name",
  TRUE ~ NA_character_
)

linkage_id_col <- dplyr::case_when(
  "ID" %in% names(linkage) ~ "ID",
  "gutsID" %in% names(linkage) ~ "gutsID",
  "GUTSID" %in% names(linkage) ~ "GUTSID",
  TRUE ~ NA_character_
)

stopifnot(!is.na(linkage_name_col))
stopifnot(!is.na(linkage_id_col))

linkage_key <- linkage %>%
  dplyr::mutate(
    Name_raw = as.character(.data[[linkage_name_col]]),
    gutsID = as.character(.data[[linkage_id_col]]),
    Name_title = make_title_key(Name_raw),
    Name_title_key = clean_name(Name_title)
  ) %>%
  dplyr::select(gutsID, Name_raw, Name_title, Name_title_key) %>%
  dplyr::distinct()

# ------------------------------------------------------------
# 4) Join IC -> linkage to assign gutsID
# ------------------------------------------------------------
ic_linked <- ic_fixed %>%
  dplyr::left_join(
    linkage_key %>%
      dplyr::select(gutsID, linkage_Name_title = Name_title, Name_title_key),
    by = "Name_title_key"
  )

# ------------------------------------------------------------
# 5) Audit checks
# ------------------------------------------------------------

# A) Which hard-coded names were applied, and did they now match linkage?
hardcode_status <- ic_linked %>%
  dplyr::filter(hardcode_applied) %>%
  dplyr::transmute(
    issue_type = dplyr::case_when(
      is.na(gutsID) | gutsID == "" ~ "hardcode_applied_still_unmatched",
      TRUE ~ "hardcode_applied_matched"
    ),
    Name_title_original,
    Name_title,
    Name_title_key,
    ResponseId,
    gutsID,
    details = dplyr::case_when(
      is.na(gutsID) | gutsID == "" ~ "hardcode applied but still no linkage match",
      TRUE ~ paste0("matched linkage name: ", linkage_Name_title)
    )
  )

# B) IC names that still do not exist in linkage
ic_unmatched <- ic_linked %>%
  dplyr::filter(is.na(gutsID) | gutsID == "") %>%
  dplyr::transmute(
    issue_type = "ic_name_missing_in_linkage",
    Name_title_original,
    Name_title,
    Name_title_key,
    ResponseId,
    gutsID = NA_character_,
    details = dplyr::case_when(
      hardcode_applied ~ "hardcode applied but still unmatched",
      TRUE ~ "no hardcode applied"
    )
  )

# C) Linkage duplicate keys
linkage_dupes <- linkage_key %>%
  dplyr::count(Name_title_key, Name_title, sort = TRUE) %>%
  dplyr::filter(n > 1) %>%
  dplyr::transmute(
    issue_type = "linkage_duplicate_name_key",
    Name_title_original = NA_character_,
    Name_title,
    Name_title_key,
    ResponseId = NA_character_,
    gutsID = NA_character_,
    details = paste0("n_rows=", n)
  )

# D) ResponseIds mapping to multiple gutsIDs
ppn_gutsid_conflicts <- ic_linked %>%
  dplyr::filter(!is.na(ResponseId) & ResponseId != "") %>%
  dplyr::group_by(ResponseId) %>%
  dplyr::summarise(
    n_gutsIDs = dplyr::n_distinct(gutsID, na.rm = TRUE),
    gutsIDs = paste(sort(unique(stats::na.omit(gutsID))), collapse = " | "),
    .groups = "drop"
  ) %>%
  dplyr::filter(n_gutsIDs > 1) %>%
  dplyr::transmute(
    issue_type = "responseid_maps_to_multiple_gutsid",
    Name_title_original = NA_character_,
    Name_title = NA_character_,
    Name_title_key = NA_character_,
    ResponseId,
    gutsID = NA_character_,
    details = paste0("gutsIDs=", gutsIDs)
  )

F_review <- dplyr::bind_rows(
  hardcode_status,
  ic_unmatched,
  linkage_dupes,
  ppn_gutsid_conflicts
)

write_audit_csv(F_review, "F_linkage_review.csv")

# ------------------------------------------------------------
# 6) Build deterministic ppn -> gutsID map
# ------------------------------------------------------------
ppn_to_gutsid <- ic_linked %>%
  dplyr::filter(!is.na(ResponseId) & ResponseId != "") %>%
  dplyr::group_by(ResponseId) %>%
  dplyr::summarise(
    gutsID = {
      vals <- sort(unique(stats::na.omit(gutsID)))
      if (length(vals) == 0) NA_character_ else vals[1]
    },
    .groups = "drop"
  ) %>%
  dplyr::rename(ppn = ResponseId)

# ------------------------------------------------------------
# 7) Propagate gutsID to Wave 0 downstream datasets
# ------------------------------------------------------------
add_gutsid <- function(df, map_tbl) {
  df %>%
    dplyr::mutate(ppn = trimws(as.character(ppn))) %>%
    dplyr::left_join(map_tbl, by = "ppn")
}

demo_linked <- add_gutsid(demo_src, ppn_to_gutsid)
survey_linked <- add_gutsid(survey_src, ppn_to_gutsid)

# ------------------------------------------------------------
# 8) Final outputs
# ------------------------------------------------------------
ic_ledger_final <- ic_linked
demo_final <- demo_linked
survey_final <- survey_linked

move_gutsid_first <- function(df) {
  if (!"gutsID" %in% names(df)) return(df)
  df %>% dplyr::select(gutsID, dplyr::everything())
}

ic_ledger_final <- move_gutsid_first(ic_ledger_final)
demo_final <- move_gutsid_first(demo_final)
survey_final <- move_gutsid_first(survey_final)

# ------------------------------------------------------------
# 9) Receipt summary
# ------------------------------------------------------------
F_receipt <- tibble::tibble(
  n_ic_rows = nrow(ic_ledger_final),
  n_demo_rows = nrow(demo_final),
  n_survey_rows = nrow(survey_final),

  n_ic_with_gutsID = sum(!is.na(ic_ledger_final$gutsID) & ic_ledger_final$gutsID != "", na.rm = TRUE),
  n_demo_with_gutsID = sum(!is.na(demo_final$gutsID) & demo_final$gutsID != "", na.rm = TRUE),
  n_survey_with_gutsID = sum(!is.na(survey_final$gutsID) & survey_final$gutsID != "", na.rm = TRUE),

  n_audit_issues = nrow(F_review),
  n_hardcode_applied = sum(ic_fixed$hardcode_applied, na.rm = TRUE),
  n_hardcode_applied_matched = sum(F_review$issue_type == "hardcode_applied_matched", na.rm = TRUE),
  n_hardcode_applied_still_unmatched = sum(F_review$issue_type == "hardcode_applied_still_unmatched", na.rm = TRUE),
  n_ic_unmatched_names = nrow(ic_unmatched),
  n_linkage_duplicate_keys = nrow(linkage_dupes),
  n_ppn_conflicts = nrow(ppn_gutsid_conflicts)
)

write_receipt_csv(F_receipt, "F_linkage_receipt.csv")

message("Audit Step F (Wave 0) complete: gutsID linked and propagated.")

```



```{r}
###############################################################################
# CHUNK 1) MEASURE OVERVIEW MANIFEST (T0) -> LOCKED REGISTRIES
#
# Purpose
#   Read the T0 behavioral overview file (ordered to match Qualtrics), clean it,
#   validate only the minimum requirements needed for downstream steps, and then
#   produce TWO canonical registries:
#
#     1) registry_items    (item-level; one row per overview row)
#     2) registry_measures (measure-level; one row per short_name x component)
#
# Philosophy
#   Treat the overview as the authoritative source of truth. We therefore avoid
#   enforcing "descriptor consistency" checks that can legitimately vary by item
#   (e.g., references, subscales).
#
# Wave 0 notes
#   - T0 is shorter than later waves
#   - T0 has no sociometrics
#   - Expected components are: sensitive, questionnaire
###############################################################################

suppressPackageStartupMessages({
  library(readxl)
  library(janitor)
  library(dplyr)
  library(stringr)
  library(tidyr)
  library(purrr)
})

# --------------------------------------------------------------------------
# STEP X.1) IMPORT OVERVIEW (T0)
# --------------------------------------------------------------------------

overview_path <- file.path(dir0, "Behavioral/T0/T0_behavioral_overview.xlsx")

if (!file.exists(overview_path)) {
  stop("T0_behavioral_overview.xlsx not found.")
}

overview_raw <- readxl::read_excel(
  path = overview_path,
  sheet = 1,
  .name_repair = "minimal"
)

overview <- overview_raw %>%
  janitor::clean_names() %>%
  dplyr::mutate(
    dplyr::across(
      .cols = where(is.character),
      .fns  = ~ stringr::str_squish(.x)
    )
  )

message("Overview rows: ", nrow(overview))
message("Overview columns: ", ncol(overview))

# --------------------------------------------------------------------------
# STEP X.2) CLEAN OVERVIEW (LOCKED)
# --------------------------------------------------------------------------

# Drop obvious Excel artifacts if present
overview <- overview %>%
  dplyr::select(-dplyr::any_of(c("b", "x", "unnamed_2")))

# Harmonize Wave 0 column names to the same registry conventions used later
overview <- overview %>%
  dplyr::rename(
    short_name = short_name,
    variable_name = variable_name,
    variable_label = variable_label,
    variable_values = variable_values,
    variable_values_label = variable_values_label,
    original_english_or_dutch = original_english_or_dutch,
    english_translation = english_translation,
    original_dutch = original_dutch,
    references = references,
    amount_of_items_in_original_questionnaire = amount_of_items_in_original_questionnaire,
    amount_of_items_in_guts_questionnaire = amount_of_items_in_guts_questionnaire
  )

# Wave 0 uses a different reversed-items column name than T1
if ("reversed_items_0_no_1_yes" %in% names(overview)) {
  overview <- overview %>%
    dplyr::rename(reversed_items = reversed_items_0_no_1_yes)
}

# Fill down implied continuity for key columns
fill_cols <- c(
  "short_name",
  "component",
  "questionnaire",
  "subscales",
  "original_english_or_dutch",
  "back_translation_if_necessary_by",
  "references",
  "amount_of_items_in_original_questionnaire",
  "amount_of_items_in_guts_questionnaire"
)

overview <- overview %>%
  tidyr::fill(dplyr::any_of(fill_cols), .direction = "down")

# String hygiene for key columns
overview <- overview %>%
  dplyr::mutate(
    dplyr::across(
      .cols = dplyr::any_of(c("short_name", "component", "variable_name")),
      .fns  = ~ stringr::str_squish(.x)
    ),
    component = dplyr::if_else(is.na(component), NA_character_, tolower(component))
  )

# Coerce flags / counts into stable types
overview <- overview %>%
  dplyr::mutate(
    reversed_items = suppressWarnings(as.integer(reversed_items)),
    amount_of_items_in_original_questionnaire =
      suppressWarnings(as.integer(amount_of_items_in_original_questionnaire)),
    amount_of_items_in_guts_questionnaire =
      suppressWarnings(as.integer(amount_of_items_in_guts_questionnaire))
  )

# --------------------------------------------------------------------------
# STEP X.3) MINIMUM VALIDATIONS (FAIL FAST ONLY WHEN NECESSARY)
# --------------------------------------------------------------------------

required_cols <- c("short_name", "component", "variable_name")
missing_required <- setdiff(required_cols, names(overview))
if (length(missing_required) > 0) {
  stop("Overview is missing required columns: ", paste(missing_required, collapse = ", "))
}

bad_short <- which(is.na(overview$short_name) | overview$short_name == "")
bad_comp  <- which(is.na(overview$component)  | overview$component  == "")
bad_var   <- which(is.na(overview$variable_name) | overview$variable_name == "")

if (length(bad_short) > 0) stop("Overview has missing short_name in row(s): ", paste(head(bad_short, 20), collapse = ", "))
if (length(bad_comp)  > 0) stop("Overview has missing component in row(s): ", paste(head(bad_comp, 20), collapse = ", "))
if (length(bad_var)   > 0) stop("Overview has missing variable_name in row(s): ", paste(head(bad_var, 20), collapse = ", "))

valid_components <- c("sensitive", "questionnaire")
unexpected_components <- overview %>%
  dplyr::distinct(component) %>%
  dplyr::filter(!component %in% valid_components)

if (nrow(unexpected_components) > 0) {
  stop(
    "Overview has unexpected component value(s): ",
    paste(unexpected_components$component, collapse = ", "),
    "\nExpected: ", paste(valid_components, collapse = ", ")
  )
}

dups_within_measure <- overview %>%
  dplyr::count(short_name, component, variable_name) %>%
  dplyr::filter(n > 1)

if (nrow(dups_within_measure) > 0) {
  print(dups_within_measure, n = 200)
  stop("Duplicate variable_name(s) detected within short_name x component. See printed table.")
}

overview_summary <- overview %>%
  dplyr::count(short_name, component, name = "n_items") %>%
  dplyr::arrange(short_name, component)

print(overview_summary)
message(
  "Manifest locked: ", nrow(overview),
  " rows across ", dplyr::n_distinct(overview$short_name),
  " measures."
)

# --------------------------------------------------------------------------
# STEP X.4) BUILD REGISTRIES
# --------------------------------------------------------------------------

parse_numeric_range <- function(x) {
  x0 <- str_squish(tolower(as.character(x)))
  if (is.na(x0) || x0 == "") return(list(ok = FALSE, min = NA_real_, max = NA_real_))
  m <- str_match(x0, "^(\\d+)\\s*to\\s*(\\d+)$")
  if (all(is.na(m))) return(list(ok = FALSE, min = NA_real_, max = NA_real_))
  list(ok = TRUE, min = as.numeric(m[2]), max = as.numeric(m[3]))
}

parse_value_labels <- function(x) {
  x0 <- as.character(x)
  if (is.na(x0) || str_squish(x0) == "") return(list(ok = FALSE, levels = NULL, n_levels = 0L))

  parts <- str_split(x0, pattern = "[;\n\r]+", simplify = FALSE)[[1]]
  parts <- str_squish(parts)
  parts <- parts[parts != ""]

  m <- str_match(parts, "^\\s*(\\d+)\\s*[:=]\\s*(.+?)\\s*$")
  if (nrow(m) == 0 || all(is.na(m[, 2]))) return(list(ok = FALSE, levels = NULL, n_levels = 0L))

  ok_idx <- which(!is.na(m[, 2]) & !is.na(m[, 3]))
  if (length(ok_idx) == 0) return(list(ok = FALSE, levels = NULL, n_levels = 0L))

  keys <- m[ok_idx, 2]
  vals <- m[ok_idx, 3]

  keep <- !duplicated(keys)
  keys <- keys[keep]
  vals <- vals[keep]

  lv <- vals
  names(lv) <- keys

  list(ok = TRUE, levels = lv, n_levels = length(lv))
}

registry_items <- overview %>%
  dplyr::mutate(
    overview_row = dplyr::row_number(),

    short_name = str_squish(as.character(short_name)),
    component = tolower(str_squish(as.character(component))),
    variable_name = str_squish(as.character(variable_name)),

    reversed_flag = suppressWarnings(as.integer(reversed_items)),
    reversed_flag = dplyr::if_else(is.na(reversed_flag), 0L, reversed_flag),

    is_prefix = str_ends(variable_name, "_"),
    variable_name_base = str_remove(variable_name, "_$"),

    range_parsed = purrr::map(variable_values, parse_numeric_range),
    scale_min = purrr::map_dbl(range_parsed, ~ .x$min),
    scale_max = purrr::map_dbl(range_parsed, ~ .x$max),
    range_ok = purrr::map_lgl(range_parsed, ~ .x$ok),

    levels_parsed = purrr::map(variable_values_label, parse_value_labels),
    levels_ok = purrr::map_lgl(levels_parsed, ~ .x$ok),
    n_levels = purrr::map_int(levels_parsed, ~ .x$n_levels),

    scale_type = dplyr::case_when(
      range_ok  ~ "numeric_range",
      levels_ok ~ "categorical_levels",
      TRUE      ~ "unknown"
    )
  ) %>%
  dplyr::arrange(overview_row)

first_non_empty <- function(x) {
  x0 <- str_squish(as.character(x))
  x0 <- x0[!is.na(x0) & x0 != ""]
  if (length(x0) == 0) NA_character_ else x0[1]
}

first_non_na_int <- function(x) {
  x0 <- suppressWarnings(as.integer(x))
  x0 <- x0[!is.na(x0)]
  if (length(x0) == 0) NA_integer_ else x0[1]
}

unique_non_empty <- function(x) {
  x0 <- str_squish(as.character(x))
  x0 <- x0[!is.na(x0) & x0 != ""]
  unique(x0)
}

registry_measures <- registry_items %>%
  dplyr::group_by(component, short_name) %>%
  dplyr::summarise(
    overview_row_start = min(overview_row),
    overview_row_end = max(overview_row),

    n_items_manifest = dplyr::n(),
    variable_names_manifest = list(variable_name),

    questionnaire =
      if ("questionnaire" %in% names(cur_data())) first_non_empty(questionnaire) else NA_character_,
    references =
      if ("references" %in% names(cur_data())) first_non_empty(references) else NA_character_,

    original_english_or_dutch =
      if ("original_english_or_dutch" %in% names(cur_data())) first_non_empty(original_english_or_dutch) else NA_character_,
    back_translation_if_necessary_by =
      if ("back_translation_if_necessary_by" %in% names(cur_data())) first_non_empty(back_translation_if_necessary_by) else NA_character_,

    amount_of_items_in_original_questionnaire =
      if ("amount_of_items_in_original_questionnaire" %in% names(cur_data())) first_non_na_int(amount_of_items_in_original_questionnaire) else NA_integer_,
    amount_of_items_in_guts_questionnaire =
      if ("amount_of_items_in_guts_questionnaire" %in% names(cur_data())) first_non_na_int(amount_of_items_in_guts_questionnaire) else NA_integer_,

    subscales_all =
      list(if ("subscales" %in% names(cur_data())) unique_non_empty(subscales) else character(0)),
    n_subscales =
      if ("subscales" %in% names(cur_data())) length(unique_non_empty(subscales)) else 0L,
    has_multiple_subscales = n_subscales > 1L,

    has_prefix_items = any(is_prefix, na.rm = TRUE),
    has_reversed_items = any(reversed_flag == 1L, na.rm = TRUE),
    has_numeric_ranges = any(range_ok, na.rm = TRUE),
    has_levels_info = any(levels_ok, na.rm = TRUE),

    .groups = "drop"
  ) %>%
  dplyr::arrange(component, short_name)

count_mismatch <- registry_measures %>%
  dplyr::filter(!is.na(amount_of_items_in_guts_questionnaire)) %>%
  dplyr::mutate(count_diff = n_items_manifest - amount_of_items_in_guts_questionnaire) %>%
  dplyr::filter(count_diff != 0)

if (nrow(count_mismatch) > 0) {
  message("NOTE: Item-count mismatches detected (manifest vs amount_of_items_in_guts_questionnaire).")
  print(
    count_mismatch %>%
      dplyr::select(component, short_name, n_items_manifest, amount_of_items_in_guts_questionnaire, count_diff),
    n = 200
  )
}

message("Registry items: ", nrow(registry_items), " rows.")
message("Registry measures: ", nrow(registry_measures), " component x short_name blocks.")

registry_measures_summary <- registry_measures %>%
  dplyr::count(component, name = "n_measures") %>%
  dplyr::arrange(component)

print(registry_measures_summary)

# Objects available downstream:
#   - overview
#   - registry_items
#   - registry_measures

```



```{r}
###############################################################################
# CHUNK X) SPLIT MEASURES (FROM REGISTRY) + REVERSE CODING (T0)
#
# Purpose
#   Build component-specific measure dataframes using registry_items, then apply
#   reverse coding using registry_items metadata (reversed_flag + numeric ranges).
#
# Outputs
#   - sensitive_measures_list_raw
#   - questionnaire_measures_list_raw
#   - sensitive_measures_list_rc
#   - questionnaire_measures_list_rc
#
# Notes
#   - Wave 0 has no sociometrics
#   - Expected components are:
#       * sensitive
#       * questionnaire
###############################################################################

suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(purrr)
  library(tidyr)
})

stopifnot(exists("registry_items"))
stopifnot(exists("demo_final"))
stopifnot(exists("survey_final"))

# --------------------------------------------------------------------------
# Helpers
# --------------------------------------------------------------------------

get_guts_col <- function(df) {
  guts_candidates <- c("gutsID", "guts_id", "gutsid", "GUTSID", "GutsID")
  guts_col <- guts_candidates[guts_candidates %in% names(df)][1]
  if (is.na(guts_col)) stop("Could not find a gutsID column in dataset.")
  guts_col
}

# Standardize registry once for downstream use
registry_items2 <- registry_items %>%
  mutate(
    component = tolower(str_squish(as.character(component))),
    short_name = str_squish(as.character(short_name)),
    variable_name = str_squish(as.character(variable_name)),
    variable_name_base = if ("variable_name_base" %in% names(.)) {
      str_squish(as.character(variable_name_base))
    } else {
      str_remove(variable_name, "_$")
    },
    is_prefix = if ("is_prefix" %in% names(.)) {
      as.logical(is_prefix)
    } else {
      str_ends(variable_name, "_")
    }
  )

# --------------------------------------------------------------------------
# Splitter (registry-driven)
# --------------------------------------------------------------------------

split_component_measures_registry <- function(df, component_name, registry_items_df) {

  guts_col <- get_guts_col(df)

  reg <- registry_items_df %>%
    filter(component == component_name) %>%
    filter(!is.na(short_name), short_name != "") %>%
    filter(!is.na(variable_name), variable_name != "")

  if (nrow(reg) == 0) {
    message("No registry rows found for component: ", component_name)
    return(list())
  }

  measures <- reg %>%
    group_by(short_name) %>%
    summarise(vars = list(variable_name), .groups = "drop") %>%
    mutate(
      data = purrr::map(vars, function(v) {

        selected_cols <- character(0)

        for (var in v) {

          prefix <- if (str_ends(var, "_")) var else paste0(var, "_")
          base   <- str_remove(var, "_$")

          exact_candidates <- unique(c(var, base))
          exact_match <- exact_candidates[exact_candidates %in% names(df)]
          prefix_match <- names(df)[str_starts(names(df), fixed(prefix))]

          if (length(exact_match) > 0 && length(prefix_match) > 0) {
            selected_cols <- c(selected_cols, exact_match[1], prefix_match)
          } else if (length(prefix_match) > 0) {
            selected_cols <- c(selected_cols, prefix_match)
          } else if (length(exact_match) > 0) {
            selected_cols <- c(selected_cols, exact_match[1])
          } else {
            stop("Variable not found in ", component_name, ": ", var)
          }
        }

        selected_cols <- unique(selected_cols)
        keep_cols <- c(guts_col, selected_cols)

        out <- df %>%
          select(all_of(keep_cols))

        if (guts_col != "gutsID") {
          out <- out %>% rename(gutsID = all_of(guts_col))
        }

        out
      })
    )

  out_list <- measures$data
  names(out_list) <- measures$short_name
  out_list
}

# --------------------------------------------------------------------------
# Split lists (RAW)
# --------------------------------------------------------------------------

sensitive_measures_list_raw <- split_component_measures_registry(
  demo_final,
  "sensitive",
  registry_items2
)

questionnaire_measures_list_raw <- split_component_measures_registry(
  survey_final,
  "questionnaire",
  registry_items2
)

print(tibble(component = "sensitive",    n_measures = length(sensitive_measures_list_raw)))
print(tibble(component = "questionnaire", n_measures = length(questionnaire_measures_list_raw)))

# --------------------------------------------------------------------------
# Reverse coding (registry-driven)
# --------------------------------------------------------------------------

apply_reverse_coding_registry <- function(measures_list_raw, component_name, registry_items_df) {

  reg <- registry_items_df %>%
    mutate(
      component = tolower(str_squish(as.character(component))),
      short_name = str_squish(as.character(short_name)),
      variable_name = str_squish(as.character(variable_name))
    ) %>%
    filter(component == component_name) %>%
    filter(reversed_flag == 1L) %>%
    filter(range_ok) %>%
    select(short_name, variable_name, scale_min, scale_max)

  if (nrow(reg) == 0) {
    message("Reverse coding (", component_name, "): no reversed items flagged with numeric ranges.")
    return(measures_list_raw)
  }

  measures_list_rc <- purrr::imap(measures_list_raw, function(df, meas) {

    info <- reg %>% filter(short_name == meas)
    if (nrow(info) == 0) return(df)

    for (i in seq_len(nrow(info))) {
      v  <- info$variable_name[i]
      mn <- info$scale_min[i]
      mx <- info$scale_max[i]

      if (v %in% names(df)) {
        x_num <- suppressWarnings(as.numeric(df[[v]]))
        df[[v]] <- ifelse(is.na(x_num), NA_real_, (mn + mx) - x_num)
        next
      }

      prefix <- if (str_ends(v, "_")) v else paste0(v, "_")
      prefixed_cols <- names(df)[str_starts(names(df), fixed(prefix))]

      if (length(prefixed_cols) > 0) {
        for (col in prefixed_cols) {
          x_num <- suppressWarnings(as.numeric(df[[col]]))
          df[[col]] <- ifelse(is.na(x_num), NA_real_, (mn + mx) - x_num)
        }
      }
    }

    df
  })

  rev_summary <- reg %>%
    group_by(short_name) %>%
    summarise(
      n_reversed = n(),
      reversed_vars = paste(variable_name, collapse = ", "),
      .groups = "drop"
    ) %>%
    arrange(desc(n_reversed))

  message("Reverse-coding summary (", component_name, "):")
  print(rev_summary, n = Inf)

  measures_list_rc
}

sensitive_measures_list_rc <- apply_reverse_coding_registry(
  sensitive_measures_list_raw,
  "sensitive",
  registry_items2
)

questionnaire_measures_list_rc <- apply_reverse_coding_registry(
  questionnaire_measures_list_raw,
  "questionnaire",
  registry_items2
)

# Objects available downstream:
#   - sensitive_measures_list_raw
#   - questionnaire_measures_list_raw
#   - sensitive_measures_list_rc
#   - questionnaire_measures_list_rc

```


```{r}
###############################################################################
# STEP X) EXPORT OUTPUTS (WAVE 0)
#
# Exports into ONE root folder:
#   <out_base_dir>/measures/
#
#   A) Measures (sensitive / questionnaire)
#      - <out_meas_dir>/<short_name>/
#
# Receipts:
#   <out_meas_dir>/_receipts/
###############################################################################

suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(purrr)
  library(readr)
  library(haven)
  library(tibble)
})

# ---- Config ----
site_id <- "lei"
ses_id  <- "00"

out_base_dir <- file.path(dir0, "Behavioral", "T0", "data_processed")
out_meas_dir <- file.path(out_base_dir, "measures")
receipts_dir <- file.path(out_meas_dir, "_receipts")

timestamp_now <- format(Sys.time(), "%Y-%m-%d %H:%M:%S")

# ---- Ensure folders exist ----
dir.create(out_meas_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(receipts_dir, recursive = TRUE, showWarnings = FALSE)

message("EXPORT ROOT (measures): ", out_meas_dir)
message("  Receipts dir:         ", receipts_dir)

# ---- Helpers ----
assert_safe_name <- function(x) {
  x <- as.character(x)
  bad <- str_detect(x, "[\\\\/:*?\"<>|\\s]")
  if (any(bad)) {
    stop(
      "Unsafe name(s) for folder/file naming (contains space or invalid path character): ",
      paste(unique(x[bad]), collapse = ", ")
    )
  }
}

# --------------------------------------------------------------------------
# A) MEASURES EXPORT -> TSV + RDS + SAV
# --------------------------------------------------------------------------

stopifnot(exists("sensitive_measures_list_rc"))
stopifnot(exists("questionnaire_measures_list_rc"))

export_sets <- list(
  sensitive    = sensitive_measures_list_rc,
  questionnaire = questionnaire_measures_list_rc
)

export_measure_set <- function(measures_list, component_name) {
  if (length(measures_list) == 0) {
    message("No measures to export for component: ", component_name)
    return(tibble())
  }

  measure_names <- names(measures_list)
  assert_safe_name(measure_names)

  purrr::imap_dfr(measures_list, function(df, short_name) {

    measure_dir <- file.path(out_meas_dir, short_name)
    dir.create(measure_dir, recursive = TRUE, showWarnings = FALSE)

    stem <- paste0("guts-", site_id, "_ses-", ses_id, "_task-", short_name)

    path_tsv <- file.path(measure_dir, paste0(stem, ".tsv"))
    path_rds <- file.path(measure_dir, paste0(stem, ".rds"))
    path_sav <- file.path(measure_dir, paste0(stem, ".sav"))

    readr::write_tsv(df, path_tsv, na = "")
    saveRDS(df, path_rds)
    haven::write_sav(df, path_sav)

    tibble(
      export_type = "measures",
      component   = component_name,
      short_name  = short_name,
      n_rows      = nrow(df),
      n_cols      = ncol(df),
      path_tsv    = path_tsv,
      path_rds    = path_rds,
      path_sav    = path_sav,
      timestamp   = timestamp_now
    )
  })
}

meas_receipt <- purrr::imap_dfr(export_sets, export_measure_set)

meas_receipt_path <- file.path(receipts_dir, "export_receipt_measures.csv")
readr::write_csv(meas_receipt, meas_receipt_path)

message("Measures export complete. Receipt: ", meas_receipt_path)

print(meas_receipt %>% count(component, name = "n_measures_exported"))
message("All exports complete. Root: ", out_meas_dir)

```

